---
title: "ESM 244 Lab 1"
author: "L. Enright"
date: "2023-01-13"
output: html_document
---

```{r setup, include=TRUE, echo = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, echo = TRUE, warning = FALSE, message = FALSE)

# first line: talking about this setup chunk (we want anyone looking at this document to see what packages we included)
# second line: talking about all other chunks. 

library(tidyverse)
library(here)
library(sf)
library(tmap)
```

### Read in the data
shortcut to new chunk: command, option, I 
read.csv <- built in function
read_csv <- improved version in the tidyverse, has added functionality! Use read_csv ! 

```{r}
sf_trees <- read_csv(here('data', 'sf_trees', 'sf_trees.csv'))

#look at lab key for metadata info

#explore the data 

top_5_status <- sf_trees %>% 
  group_by(legal_status) %>%
  summarize(tree_count = n()) %>%
  slice_max(tree_count, n = 5) %>% #says find the max values for this variable and only keep X number (n = "")
  arrange(-tree_count) #this makes it go in order from greatest to least 
```

#Explore the Data
(in the console)

- summary() :can be helpful to look at max/min long and lat -- can give you an idea if you geo data might be funky (i.e. trees in the pacific ocean)

- sf_trees %>% group_by(legal_status) %>% summarize(tree_count = n())

```{r}
ggplot(data = top_5_status, aes(x = fct_reorder(legal_status, tree_count),
                                y = tree_count)) +
  geom_col() +
  labs(y = "Tree count", x = "Legal Status") +
  coord_flip() + #says swap my axes
  theme_minimal()
```

